---
title: "Deforestation"
author: "Sophia Boyd"
date: "5/9/2022"
output:
  html_document:
    toc: TRUE
    theme: sandstone
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.path = "../Output/")
```


# Introduction

Much of the world's surface is covered by forest alone. 1/3 of our habitable land area is taken up by forests.

In order to determine the net change in forest cover, we need to look at how much deforestation occurs.

Due to the expansion of agriculture, the global forest area has drastically reduced. Over 12 million acres of forest is lost every year, which is driven mostly by agriculture. 

We will use the global deforestation data to further examine:

- Where are we losing forests?
- What is impacting the loss of forests?


## Load Libraries
```{r}
library(tidyverse)
library(here)
library(ggplot2)
library(dplyr)
library(maps)
library(mapdata)
library(mapproj)
library(gganimate)
library(hrbrthemes)
library(beyonce)
library(flair)
library(RColorBrewer)
library(LaCroixColoR)
library(devtools)
```

## Load Data
```{r}
forest <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/forest.csv')
write_csv(forest, here('Independent_Project',"Data","forest.csv"))
forest_area <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/forest_area.csv')
write_csv(forest_area, here('Independent_Project', "Data", "forestarea.csv"))
brazil_loss <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/brazil_loss.csv')
write_csv(brazil_loss, here('Independent_Project', "Data", "brazilloss.csv"))
soybean_use <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/soybean_use.csv')
write_csv(soybean_use, here('Independent_Project', "Data", "soybeanuse.csv"))
vegetable_oil <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-04-06/vegetable_oil.csv')
write_csv(vegetable_oil, here('Independent_Project', "Data", "vegetableoil.csv"))
```


## Visuals {.tabset}

### Global Forest Area

Forests take up 38% of the habitable land area which means they are spread throughout the entire world
```{r}
newforest_area <- forest_area %>%
  rename(country = entity) %>% 
  mutate_at("country", str_replace, "Americas", "USA") %>% 
  select(-c(code)) %>% 
  filter(year %in% c("2020"))

country <- map_data("world")

country_map <- merge(country, newforest_area, by.x = "region", by.y = "country") 

country_map <- arrange(country_map, group, order)

ggplot(country_map, aes(x = long, y = lat, group = group, fill = forest_area)) +
  geom_polygon(color = "black") +
  theme_minimal() +
  scale_fill_distiller(palette = "Dark2") +
  coord_map(projection = "mercator",
            xlim = c(-180,180)) +
  theme(panel.background = element_rect(fill = "lightblue")) +
  labs(fill = "Amount of Forest Area")

```

### Loss of Brazilian Forest

There are several different drivers that has caused Brazil to lose much of its Brazilian forest over the years.

```{r}
#Create a new data set
brazil_long <- brazil_loss %>% 
  pivot_longer(cols = commercial_crops:small_scale_clearing,
               names_to = "Variables",
               values_to = "Values")

brazil_pie <- brazil_long %>% 
  arrange(desc(Variables)) %>% 
  mutate(prop = Values / sum(brazil_long$Values) *100) %>% 
  mutate(ypos = cumsum(prop) - 0.5*prop)

```

```{r}
#Loss in 2001
brazil_2001 <- brazil_pie %>% 
  filter(year %in% c("2001"))

ggplot(brazil_2001, aes(x = "", y = prop, fill = Variables))+
  geom_bar(stat = "identity", width = 1, color = "white")+
  coord_polar("y", start = 0)+
  theme_void() +
  scale_fill_manual(values = lacroix_palette(type = "paired"))

#Loss in 2013
brazil_2013 <- brazil_pie %>% 
  filter(year %in% c("2013"))

ggplot(brazil_2013, aes(x = "", y = prop, fill = Variables))+
  geom_bar(stat = "identity", width = 1, color = "white")+
  coord_polar("y", start = 0)+
  theme_void()+
  scale_fill_manual(values = lacroix_palette(type = "paired"))
```

  
### Soybean vs. Vegetable Oil Produced

The top drivers of global deforestation consist of vegetable oil crops, soybean and palm oil. 

Over the past decades, the rising demand for these foods has began threatening ecosystem functions in the world's forest areas.
```{r}
#Soybean Production and Use

soybean_new <- soybean_use %>% 
  pivot_longer(cols = human_food:processed,
               names_to = "Variables",
               values_to = "Values") %>% 
  select(-c(code)) %>% 
  drop_na() %>% 
  filter(year > "1999",
         entity %in% c("Americas", "Russia", "Brazil", "Germany"))
  

 soybean_new %>% 
  ggplot(aes(x = year, y = Values, color = Variables)) +
  geom_line() + 
  geom_point() + 
  facet_wrap(~entity, scales = "free") +
  theme_ipsum() + 
  labs(title = "Production of soybean and usage",
       x = "Year",
       y = "Soybean Usage",
       color = "Food used for") +
  scale_color_manual(values = lacroix_palette("PassionFruit", type = "discrete")) +
  transition_reveal(year)


#Vegetable Oil Production 

vegetable_new <- vegetable_oil %>% 
  select(-c(code)) %>% 
  drop_na() %>% 
  filter(crop_oil %in% c("Soybean", "Maize", "Sunflower", "Olive, virgin", "Palm"),
         year > "1999",
         entity %in% c("Americas", "Russia", "Brazil", "Germany"))

vegetable_new %>% 
  ggplot(aes(x = year, y = production, color = crop_oil)) +
  geom_line() + 
  geom_point() + 
  facet_wrap(~entity, scales = "free") +
  theme_ipsum() + 
  labs(title = "Production of vegetable oil by crop type",
       x = "Year",
       y = "Oil production (tonnes)",
       color = "Crop Type") +
  scale_color_manual(values = lacroix_palette("PassionFruit", type = "discrete")) +
  transition_reveal(year)
```